Liveness-Based Garbage Collection for Lazy Languages

### 2015-03-21 paper/paper.tex
Added generating liveness equations section

@@

>>>>

  Consider the example program in Figure~\ref{fig:mot-example2}.

@@

\section{Generating Liveness Equations}
\begin{figure*}[t!]
\begin{picture}(100,130)(0,-90)
  \begin{tabular}{cc}
    \begin{boxedminipage}{.5\textwidth}
      {\sf
        \renewcommand{\arraystretch}{1}{
          \begin{uprogram}
            \UNL{1} (\DEFINE\ (\length~\xl)
            \UNL{2}  $\pi_1\!\!:\, $(\LET\ \px\ $\leftarrow $\ (\NULLQ~\xl) \IN
            \UNL{3} \hspace*{.05cm} $\pi_2\!\!:\,$(\SIF\ $\pi_3\!\!:\,$ \px
            \UNL{4} \hspace*{1.15cm} $\pi_4\!\!:\, (\RETURN_1~0)$
            \UNL{4} \hspace*{1.15cm}    $\pi_5\!\!:\, $(\LET~\pu\  $\leftarrow$  (\CDR~\xl)  \IN
            \UNL{5} \hspace*{1.15cm}   $\pi_6\!\!:\, $(\LET~\py\  $\leftarrow$  (\length~\pu)  \IN
            \UNL{6} \hspace*{1.15cm} $\pi_7\!\!:\, (\RETURN_2~(1~+~\py))$)))))
        \end{uprogram}}
        \renewcommand{\arraystretch}{1}{
	  \begin{uprogram}
	  \UNL{1} $\pi_\mainpgm\!\!:\, $(\LET\  \pa\  $\leftarrow$ \ldots  \IN
	  \UNL{2}   \hspace*{.78cm}              (\LET\ \pb\  $\leftarrow$  \ldots \IN 
          \UNL{3}   \hspace*{.78cm}    $\pi_8\!\!:\, $(\LET\ \pw\  $\leftarrow$  (\length\ \py) \IN~$\pi_{9}\!\!:\,$ (\RETURN\ \pw)))))))
\end{uprogram}}
        } 
    \end{boxedminipage}
    & 
    \begin{boxedminipage}{.5\textwidth}
      {\sf
        \begin{tabular}{lc}
        $\cal{M} = $ 
            \begin{tabular}{lc}            
              $[$
              $\pi_{\mbox{7}} \mapsto \lbrace \py.\epsilon, \pu.\Lf{\length}{\mbox{1}}{\epsilon}, \
              \xl.\mbox{1}\Lf{\length}{\acdr}{\epsilon}, \xl.\epsilon  \rbrace ,$ \\ 
              $\pi_{\mbox{4}} \mapsto \lbrace \rbrace , $\\
              $\pi_{\mbox{3}} \mapsto \lbrace \px.\epsilon, \xl.\epsilon \rbrace $  \\
              $]$
            \end{tabular} \\
            $\cal{P} = [$
              \begin{tabular}{lc}
              $\pi_{\mbox{1}} \mapsto \lbrace \pi_{\mbox{3}},\pi_{\mbox{4}}, \pi_{\mbox{7}} \rbrace, $ \
              $ \pi_{\mbox{2}} \mapsto \lbrace \pi_{\mbox{3}},\pi_{\mbox{4}}, \pi_{\mbox{7}} \rbrace, $ \\
              $ \pi_{\mbox{3}} \mapsto \lbrace \pi_{\mbox{3}} \rbrace, $ \
              $ \pi_{\mbox{4}} \mapsto \lbrace \pi_{\mbox{4}} \rbrace, $ \
              $ \pi_{\mbox{5}} \mapsto \lbrace \pi_{\mbox{7}} \rbrace, $ \\
              $ \pi_{\mbox{6}} \mapsto \lbrace \pi_{\mbox{7}} \rbrace, $ \
              $ \pi_{\mbox{7}} \mapsto \lbrace \pi_{\mbox{7}} \rbrace $ \\
              $]$
            \end{tabular}
        \end{tabular}
      }
    \end{boxedminipage}
    \\
    (a) Example program.&
    \renewcommand{\arraystretch}{.9}{\begin{tabular}[t]{ll}
        (b) {\em e-path} to liveness environment map \\
        (c) Prog pt to {\em e-paths } map
    \end{tabular}}
  \end{tabular}
\end{picture}
\kern -3ex

\caption{Example Program and its liveness maps.}\label{fig:mot-example2}    
\end{figure*}

 For our example program which has  a
single function \length, this generates the following equations:

  \begin{eqnarray*}
    \Lf{\length}{1}{\sigma}
    &=& \epsilonset \cup \acdr\Lf{\length}{1}{\epsilon}
   \end{eqnarray*}

\subsection{Generating Liveness Equations for Function bodies}
For the running example, $\length$ has calls from $\mainpgm$ at $\pi_8$
and a recursive call at $\pi_6$.
\newcommand{\deltacall}[3]{\delta_{#1}({#2},{#3})}
So $\sigma_{\length} =
     \deltacall{\length}{\pi_8}{\mainpgm}  \cup \deltacall{\length}{\pi_6}{\length}$.
Filling in  the values gives:

\begin{align*}
\sigma_{\length}    &=
 \sigma_{\!all}  ~\cup~ \acdr\Lf{length}{1}{\sigma_{\mathit{\length}}}
\end{align*}

For the running example, containing unique GC points (corresponding to unique {\em e-paths} sets) $\pi_3, \pi_4, \pi_7$ in $\length$ and
$\pi_9$ in $e_\mainpgm$, this gives

\begin{align*}
\Lanv{7}{\xl} &= \acdr\Lf{\length}{1}{\epsilon} \cup \epsilonset \\
\Lanv{7}{\pu} &= \Lf{\length}{1}{\epsilon} \\
\Lanv{7}{\py} &= \epsilonset \\
\Lanv{3}{\xl} &= \epsilonset \\
\Lanv{3}{\px} &= \epsilonset 
\\[1ex]
  \Lanv{9}{\pb}  &= \Lf{\length}{1}{\sigma_{\mathit{\!all}}} \\
  \Lanv{9}{\pw}  &= \sigma_{\mathit{\!all}}
\end{align*}

\subsection{Solving liveness equations---the grammar interpretation}
The  equations above can now be re-interpreted as a
context-free grammar (CFG)  on the alphabet $\lbrace\acar, \acdr,
\bcar, \bcdr\rbrace$.  Let  \var{$X$} denote the non-terminal for
a variable  $X$ occurring on  the LHS of the  equations generated
from the analysis.  We can  think of the resulting productions as
being associated with several  grammars, one for each non-terminal
\var{\Lanv{i}{x}} regarded as a start symbol.  As an example, the
grammar    for   \var{\Lanv{7}{\xl}}   comprises    the   following
productions:
 \begin{eqnarray*}
\var{\Lanv{7}{\xl}}  &\rightarrow&   \var{\acdr \Lf
  {\length}{1}{\epsilon}} \mid \epsilon \\
\end{eqnarray*}
